Generalized Tractability

### 2023-04-03 GeneralizedTractability_230310.tex
Fred's Home Computer Thm 3 optimality
Updates from Overleaf

Also, if  $\vp^* \in \cp_{\textup{opt}}$, then \eqref{eq:tract_def} must not hold for all $\vp \in [\vzero, \vp^*)$.  By the sufficiency argument, this implies that \eqref{eq:tractiff4} cannot hold for all $\vp \in [\vp^*/2)$.  This is the second condition for membership in $\widetilde{\cp}_{\textup{opt}}$.

@@

\peter{
\section{Weak $T$-$(s,t)$-weak tractability (work in progress)} \label{sec:weak}

@@

We show a necessary and sufficient condition for $T$-$(s,t)$-weak tractability, working analogously to \cite{WW17}.
\begin{theorem}\label{thm:weak}
For $c>0$, define
\[
\sigma (c,d,s,\vp):= \sum_{i=1}^\infty \exp \left(-c\, T(\lambda_{i,d}^{-s},1,\vp)\right).
\]
Then the problem is $T$-$(s,t)$-weakly tractable if and only if
\[
\mu (c,s,t,\vp)=\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,p,q)\right)<\infty
\]
holds for all $c>0$.
\end{theorem}
\begin{proof}
 We first show the necessity of the condition in the theorem.   Indeed, suppose that the problem is  $T$-$(s,t)$-weakly tractable. Then, for any $c>0$ there exists a positive integer $V=V(c)$ such that  
 \[
 n(\varepsilon,d)\le 
 \left\lfloor \exp\left(c\,
 \left(T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)\right)
 \right) \right\rfloor
 \]
 whenever $\varepsilon^{-1} + d \ge V$. For $d\in\naturals$ 
 choose $\varepsilon>0$ such that $\varepsilon^{-1}\ge \max\{1,V-d\}$. Since the eigenvalues $\lambda{i,d}$ are non-increasing in $i$, we know that
 \[
 \lambda_{\lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1}^2 \le \varepsilon^2
 \]
 for any $\varepsilon$ satisfying $\varepsilon^{-1}\ge \max\{1,V-d\}$.
Let 
\[
n=n(\varepsilon):= \lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1
\]
and
\[
n^*:= \lfloor \exp(c\,
 (T(\max\{1,(V-d)^s\},1,\vp) + T(0,d^t,\vp))) \rfloor+1.
\]
By varying $\varepsilon^{-1}$ ini the interval 
$[\max\{1,V-d\},\infty)$ we can make $n(\varepsilon)$ attain 
any integer value greater than or equal to $n^*$.

[TO BE CONTINUED]

>>>>

Also, if  $\vp^* \in \cp_{\textup{opt}}$, then \eqref{eq:tract_def} must not hold for all $\vp \in [\vzero, \vp^*)$.  By the sufficiency argument, this implies that \eqref{eq:tractiff4} cannot hold for all $\vp \in [\vzero,\vp^*/2)$.  This is the second condition for membership in $\widetilde{\cp}_{\textup{opt}}$.

@@

\peter{
\section{Weak $T$-$(s,t)$-weak tractability} \label{sec:weak}

@@

We show a necessary and sufficient condition for $T$-$(s,t)$-weak tractability, working analogously to \cite{WW17}.
\begin{theorem}\label{thm:weak}
For $c>0$, define
\[
\sigma (c,d,s,\vp):= \sum_{i=1}^\infty \exp \left(-c\, T(\lambda_{i,d}^{-s},1,\vp)\right).
\]
Then the problem is $T$-$(s,t)$-weakly tractable if and only if
\[
\mu (c,s,t,\vp)=\sup_{d\in\naturals} \sigma (c,d,s,\vp) \exp\left(-c\, T(0,d^t,\vp)\right)<\infty
\]
holds for all $c>0$.
\end{theorem}
\begin{proof}
 We first show the necessity of the condition in the theorem.   Indeed, suppose that the problem is  $T$-$(s,t)$-weakly tractable. Then, for any $c>0$ there exists a positive integer $V=V(c)$ such that  
 \[
 n(\varepsilon,d)\le 
 \left\lfloor \exp\left(c\,
 \left(T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)\right)
 \right) \right\rfloor
 \]
 whenever $\varepsilon^{-1} + d \ge V$. For $d\in\naturals$ 
 choose $\varepsilon>0$ such that $\varepsilon^{-1}\ge \max\{1,V-d\}$. Since the eigenvalues $\lambda{i,d}$ are non-increasing in $i$, we know that
 \[
 \lambda_{\lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1}^2 \le \varepsilon^2
 \]
 for any $\varepsilon$ satisfying $\varepsilon^{-1}\ge \max\{1,V-d\}$.
Let 
\[
n=n(\varepsilon):= \lfloor \exp(c\,
 (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))) \rfloor+1
\]
and
\[
n^*:= \lfloor \exp(c\,
 (T(\max\{1,(V-d)^s\},1,\vp) + T(0,d^t,\vp))) \rfloor+1.
\]
By varying $\varepsilon^{-1}$ ini the interval 
$[\max\{1,V-d\},\infty)$ we can make $n(\varepsilon)$ attain 
any integer value greater than or equal to $n^*$.

Furthermore, we have, as $\lambda_{n,d}\le \varepsilon$ and as $T$ is monotonically increasing, 
\begin{eqnarray*}
n(\varepsilon)-1 & \le & \exp(c\, (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)))\\
& \le & \exp(c\, (T(\lambda_{n(\varepsilon),d}^{-s},1,\vp) + T(0,d^t,\vp))),
\end{eqnarray*}
which is equivalent to 
\[
\exp(-2c\, (T(\lambda_{n(\varepsilon),d}^{-s},1,\vp) + T(0,d^t,\vp))) \le \frac{1}{(n(\varepsilon)-1)^2}.
\]
Therefore,
\begin{eqnarray*}
    \exp(-2c\, T(0,d^t,\vp)) \sum_{i=1}^\infty \exp(-2c\, T(\lambda_{i,d}^{-s},1,\vp))
    &\le& \exp(-2c\, T(0,d^t,\vp))\, (n^*-1) + \sum_{i=n^*}^\infty \frac{1}{(i-1)^2}\\
    &\le& \exp(-2c\, T(0,d^t,\vp))\, (n^*-1) + \frac{\pi^2}{6}.
\end{eqnarray*}
Moreover,
\begin{eqnarray*}
    \exp(-2c\, T(0,d^t,\vp))\, (n^*-1) &\le & \exp(-2c\, T(0,d^t,\vp))\, 
    \exp (c\, T(V^s,1,\vp) + c\, T(0,d^t,\vp))\\
    &\le & \exp (c\, T(V^s,1,\vp))\\
    &<& \infty,
\end{eqnarray*}
independently of $d$. This implies that 
\[
\exp(-2c\, T(0,d^t,\vp)) \sum_{i=1}^\infty \exp(-2c\, T(\lambda_{i,d}^{-s},1,\vp))<\infty,
\]
also independently of $d$. As $c>0$ was chosen arbitrarily, we can replace $2c$ by $c$ and see
necessity of the condition in the theorem.

We now show sufficiency.

Suppose that $\mu(c,s,t,\vp) <\infty$ for all $c>0$. It is no loss of generality to assume that 
$\mu(c,s,t,\vp) \ge 1$. Since the terms of the series in the definition of $\mu(c,s,t,\vp) $ are non-increasing, we have 
\[
\exp (-c\, T(0,d^t,\vp))\, n\, \exp (-c\, T(\lambda_{n,d}^{-s},1,\vp)) \le \mu (c,s,t,\vp).
\]
Then we obtain 
\[
\exp (-c\, T(\lambda_{n,d}^{-s},1,\vp)) \le \frac{1}{n}\, \mu (c,s,t,\vp)\, 
\exp (c\, T(0,d^t,\vp)),
\]
which is equivalent to 
\[
\exp (c\, T(\lambda_{n,d}^{-s},1,\vp)) \ge \frac{n}{\mu (c,s,t,\vp)\, 
\exp (c\, T(0,d^t,\vp))},
\]
which in turn is equivalent to 
\[
T(\lambda_{n,d}^{-s},1,\vp) \ge \frac{1}{c}\log \left(\frac{n}{\mu (c,s,t,\vp)\, 
\exp (c\, T(0,d^t,\vp))}\right)
\]
as long as 
\[
n> \mu (c,s,t,\vp)\, \exp (c\, T(0,d^t,\vp)).
\]
This implies that for the special choice 
\[
n= \left\lceil 
   \mu (c,s,t,\vp)\, \exp \left(c\, (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))\right)
   \right\rceil 
\]
we have
\[
T(\lambda_{n,d}^{-s},1,\vp) \ge T(\varepsilon^{-1},1,\vp),
\]
which implies that $\lambda_{n,d} \le \varepsilon$, and thus $n(\varepsilon,d)\le n$. Note that the 
argument $x$ of the ceiling function in the definition of our special $n$ is at least one, which implies 
$\lceil x \rceil \le 2x$.

Consequently,
\[
n(\varepsilon, d) \le \exp\left(c\, \left(
\frac{\log (2\mu (c,s,t,\vp) )}{c}
T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp)
\right)\right).
\]
Clearly, there must exist a positive $V=V(c,s,t,\vp)$ such that when $\varepsilon^{-1} +d \ge V$ it holds true that
\[
  T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp) \ge \frac{\log (2\mu (c,s,t,\vp) )}{c}.
\]
Hence we finally have 
\[
n(\varepsilon,d) \le \exp (2c\, (T(\varepsilon^{-s},1,\vp) + T(0,d^t,\vp))).
\]
Since the latter inequality holds for every $c>0$, we have $T$-$(s,t)$-weak tractability.
